---
title: "Stat Comp HW#4"
author: "Jason Rights"
date: "October 24, 2015"
output: pdf_document
---

Problem 1:  

```{r}
secant <- function(fun, x0, x1, tol=1e-7, iter=100){
	for(i in 1:iter) {
		x2 <- x1-fun(x1)*(x1-x0)/(fun(x1)-fun(x0))
		if (abs(fun(x2)) < tol)
			return(x2)
		x0 <- x1
		x1 <- x2
	}
	stop('method did not converge')
}

f <- function(x) cos(x) - x

start.time.sec <- proc.time()
secant(f, x0=1, x1=2)
total.time.sec <- proc.time() - start.time.sec
print(total.time.sec)

newton <- function(x, f, fp, tol=1e-7, iter=100){
  i <- 1
  while(abs(eval(f)) > tol & i < iter) {
    x <- x - eval(f)/eval(D(f, 'x'))
    i <- i + 1
  }
  if(i == iter) {
    stop('method did not converge')
  }
  x
}

start.time.new <- proc.time()
newton(1, expression(cos(x)-x))
total.time.new <- proc.time() - start.time.new
print(total.time.new)

total.time.new - total.time.sec
```

The secant is typically faster, e.g., by the above amount.     

Problem 2:  

```{r}

craps <- function(ntrials){
  for(i in seq(ntrials)){
nloop <- 1
win <- 0
repeat{
  x <- sum(ceiling(6*runif(2)))
  if (nloop==1 && (x==7 | x == 11)) win <- 1
  if (nloop==1) roll <- x
  if (nloop > 1 && x == roll) win <- 1
  if (x==7 | x == 11 | win==1) break;
  nloop <- nloop +1
}
 if (i==1 && win==0) results <- 'LOSE :('
 if (i==1 && win==1) results <- 'WIN :D'
    if (i>1 && win==0) results <- c(results,'LOSE :(')
    if (i>1 && win==1) results <- c(results,'WIN :D')
  }
  return(results)
}


  set.seed(100)
  craps(3)

for(i in 1:1000){
set.seed(i)
y <- craps(10)
if (identical(y,rep("WIN :D",10))==TRUE) print(i)
}

set.seed(880)
craps(10)
```


Problem 3:


```{r}

setwd("C:/Users/Jason/Documents")
# path: directory path to input files
# file: name of the output file; it should be written to path
# nTeams: number of teams in league
# cap: money available to each team
# posReq: number of starters for each position
# points: point allocation for each category
ffvalues <- function(path, file='outfile.csv', nTeams=12, cap=200, posReq=c(qb=1, rb=2, wr=3, te=1, k=1),
                     points=c(fg=4, xpt=1, pass_yds=1/25, pass_tds=4, pass_ints=-2,
                              rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6)) {
  ## read in CSV files
k <- read.csv(file.path('proj_k15.csv'), header=TRUE, stringsAsFactors=FALSE)
qb <- read.csv(file.path('proj_qb15.csv'), header=TRUE, stringsAsFactors=FALSE)
rb <- read.csv(file.path('proj_rb15.csv'), header=TRUE, stringsAsFactors=FALSE)
te <- read.csv(file.path('proj_te15.csv'), header=TRUE, stringsAsFactors=FALSE)
wr <- read.csv(file.path('proj_wr15.csv'), header=TRUE, stringsAsFactors=FALSE)
# generate unique list of column names
cols <- unique(c(names(k), names(qb), names(rb), names(te), names(wr)))
# create a new column in each data.frame
# values are recylcled
# concept: ?Extract
k[,'pos'] <- 'k'
qb[,'pos'] <- 'qb'
rb[,'pos'] <- 'rb'
te[,'pos'] <- 'te'
wr[,'pos'] <- 'wr'

# append 'pos' to unique column list
cols <- c(cols, 'pos')

# create common columns in each data.frame
# initialize values to zero
k[,setdiff(cols, names(k))] <- 0
qb[,setdiff(cols, names(qb))] <- 0
rb[,setdiff(cols, names(rb))] <- 0
te[,setdiff(cols, names(te))] <- 0
wr[,setdiff(cols, names(wr))] <- 0

# combine data.frames by row, using consistent column order
x <- rbind(k[,cols], qb[,cols], rb[,cols], te[,cols], wr[,cols])

# calculate new columns
# convert NFL stat to fantasy points
x[,'p_fg'] <- x[,'fg']*points["fg"]
x[,'p_xpt'] <- x[,'xpt']*points["xpt"]
x[,'p_pass_yds'] <- x[,'pass_yds']*points["pass_yds"]
x[,'p_pass_tds'] <- x[,'pass_tds']*points["pass_tds"]
x[,'p_pass_ints'] <- x[,'pass_ints']*points["pass_ints"]
x[,'p_rush_yds'] <- x[,'rush_yds']*points["rush_yds"]
x[,'p_rush_tds'] <- x[,'rush_tds']*points["rush_tds"]
x[,'p_fumbles'] <- x[,'fumbles']*points["fumbles"]
x[,'p_rec_yds'] <- x[,'rec_yds']*points["rec_yds"]
x[,'p_rec_tds'] <- x[,'rec_tds']*points["rec_tds"]

# sum selected column values for every row
# this is total fantasy points for each player
x[,'points'] <- rowSums(x[,grep("^p_", names(x))])


  ## calculate dollar values

# create new data.frame ordered by points descendingly
x2 <- x[order(x[,'points'], decreasing=TRUE),]

# determine the row indeces for each position
k.ix <- which(x2[,'pos']=='k')
qb.ix <- which(x2[,'pos']=='qb')
rb.ix <- which(x2[,'pos']=='rb')
te.ix <- which(x2[,'pos']=='te')
wr.ix <- which(x2[,'pos']=='wr')

# calculate marginal points by subtracting "baseline" player's points
x2[k.ix, 'marg'] <- x2[k.ix,'points'] - x2[k.ix[12],'points']
x2[qb.ix, 'marg'] <- x2[qb.ix,'points'] - x2[qb.ix[12],'points']
x2[rb.ix, 'marg'] <- x2[rb.ix,'points'] - x2[rb.ix[24],'points']
x2[te.ix, 'marg'] <- x2[te.ix,'points'] - x2[te.ix[12],'points']
x2[wr.ix, 'marg'] <- x2[wr.ix,'points'] - x2[wr.ix[36],'points']

# create a new data.frame subset by non-negative marginal points
x3 <- x2[x2[,'marg'] >= 0,]

# re-order by marginal points
x3 <- x3[order(x3[,'marg'], decreasing=TRUE),]

# reset the row names
rownames(x3) <- NULL

# calculation for player value
x3[,'value'] <- x3[,'marg']*(12*200-nrow(x3))/sum(x3[,'marg']) + 1

# create a data.frame with more interesting columns
x4 <- x3[,c('PlayerName','pos','points','marg','value')]


  ## save dollar values as CSV file
write.csv(file=file,x4)

  ## return data.frame with dollar values
return(x4)
}



x1 <- ffvalues('.')
x1[which(x1$value > 20),]
dim(x1[which(x1$value > 20),])
```

40 players are worth more than 20 dollars.  

```{r}
setwd("C:/Users/Jason/Documents")
x1[which(x1$pos=="rb"),][15,]
```

Melvin Gordon is the 15th most valuable running back.  


```{r}
setwd("C:/Users/Jason/Documents")
x2 <- ffvalues(getwd(), '16team.csv', nTeams=16, cap=150)
x2[which(x2$value > 20),]
dim(x2[which(x2$value > 20),])
```

40 players are worth more than 20 dollars.  

```{r}
setwd("C:/Users/Jason/Documents")
temp <- x2[1:40,]
temp[which(temp$pos=="wr"),]
dim(temp[which(temp$pos=="wr"),])
```

14 wide receivers are in the top 40.  

```{r}
setwd("C:/Users/Jason/Documents")
x3 <- ffvalues('.', 'qbheavy.csv', posReq=c(qb=2, rb=2, wr=3, te=1, k=0),
        points=c(fg=0, xpt=0, pass_yds=1/25, pass_tds=6, pass_ints=-2,
                rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6))

x3[which(x3$value > 20),]
dim(x3[which(x3$value > 20),])
```

40 players are worth more than 20 dollars.  

```{r}
setwd("C:/Users/Jason/Documents")
temp <- x3[1:30,]
temp[which(temp$pos=="qb"),]
dim(temp[which(temp$pos=="qb"),])

```

5 quarterbacks are in the top 40.  



Problem 4:  

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)

n.args <- length(as.list(args(names(funs)[1])))-1
for(i in 2:length(funs)){
 n.args <- c(n.args,length(as.list(args(names(funs)[i])))-1)
}
n.args <- matrix(n.args)
colnames(n.args) <- "numb.args"
n.args <- cbind(c(1:length(n.args)),n.args) 
funs[n.args[which(n.args[,2]==max(n.args[,2])),][1]]
```

Scan has the largest number of arguments (22).  

```{r}
length(n.args[which(n.args[,2] < 1),])

```

146 do not have arguments.  



