---
title: "Stat Comp Hw#2 Jason Rights"
output: html_document
---

*(informally) Due Thursday, 17 September, 1:00 PM*

50 points total.

This assignment won't be submitted until we've covered Rmarkdown.
Create R chunks for each question and insert your R code appropriately.
Check your output by using the `Knit PDF` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
    
```{r}
cancer.df <- read.csv("https://raw.githubusercontent.com/fonnesbeck/Bios6301/master/datasets/cancer.csv")
```

    2. Determine the number of rows and columns in the data frame. (2)

```{r}
dim(cancer.df) 
```

42120 rows, 8 columns.  

    3. Extract the names of the columns in `cancer.df`. (2)

```{r}
names(cancer.df) 
```

    4. Report the value of the 3000th row in column 6. (2)

```{r}
cancer.df[3000,6] 
```

    5. Report the contents of the 172nd row. (2)

```{r}
cancer.df[172,] 
```

    6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)

```{r}
cancer.df.new <- cbind(cancer.df,(cancer.df$incidence/cancer.df$population)*10000)
colnames(cancer.df.new) <- c(colnames(cancer.df),"incidencerate")
```

    7. How many subgroups (rows) have a zero incidence rate? (2)
    
```{r}
attach(cancer.df.new)
dim(cancer.df.new[which(incidencerate==0),])
```

There are 23191 subgroups with a zero rate.  

    8. Find the subgroup with the highest incidence rate.(3)
    
```{r}
cancer.df.new[which(incidencerate==max(incidencerate)),]
```

2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)
    
            max(x)
            sort(x)
            sum(x)

```{r}
x <- c("5","12","7")
max(x)
sort(x)
#sum(x)
```

Sum produces an error message because this is a character vector and it does not make sense to sum characters. Using "max" yeilds the entry that comes last when arranged in  alphanumeric order, which in this case is 7 (as 7 comes after 5 or 1). Using sort puts them in alphanumeric order (12, 5, 7). 

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
            y[2] + y[3]

```{r}
y <- c("5",7,12)
#y[2] + y[3]
```

Y is a character vector because the character type is less flexible than the numeric type (thus, combining a character with numeric entries will make the whole vector character type). Since the "7" and "12" are characters, it does not make sense to add them, hence the error. 


    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
            z[1,2] + z[1,3]
            
```{r}
z <- data.frame(z1="5",z2=7,z3=12)
z[1,2] + z[1,3]
```

When you create z as a dataframe, it allows a combination of character and numeric entries, so the 7 and 12 are not converted to character like they are for vectors. Thus, it is fine to add them.  

3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$

```{r}
c(seq(8),c(7:1))
```    
    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$

```{r}
x <- 1
for(i in 2:5){
  x <- c(x,replicate(i,i))
}
x
```  

    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$

```{r}
x <- matrix(1,3,3)
diag(x) <- 0 
x
``` 

    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$

```{r}
x <- matrix(NA,5,4)
for(i in seq(5)){
  for(j in seq(4)){
     x[i,j] <- j^(i)
  }
}
x
``` 

4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)

```{r}
forloopforfour <- function(x,n){
  hxn <- 1
  for(i in seq(n)){
    hxn <- c(hxn,x^i)
  }
  print(sum(hxn))
}
``` 

    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)
    
```{r}
sum.natnumb <- function(mult1,mult2,N){
x <- seq(1:(N-1))
x.mult1 <- x[x%%mult1==0]
x.mult2 <- x[x%%mult2==0]
x <- c(x.mult1,x.mult2)
#print(x)
return(sum(as.numeric(unique(x))))
}
``` 

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
        
```{r}
sum.natnumb(3,5,1000)
``` 

        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
        
```{r}
sum.natnumb(4,7,1000000)
``` 

    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])

```{r}
x <- c(1,2)
  while(length(x[x%%2==0]) < 15){
    x <- c(x,sum(x[length(x)],x[length(x)-1]))
  }
print(sum(x[x%%2==0]))
    
``` 

Some problems taken or inspired by projecteuler.

[euler1]:https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2

### Notes/Grade
```{r}
# Another way to do 3.2
rep(1:5, times=1:5)

# Another way to do 3.3
+!diag(3)

# Another way to do 3.4
t(outer(1:4, 1:5, "^"))

```


Grade:
50 points + 5 bonus points = 55/50 points